---
title: "Post Thesis Defense"
output: html_document
---
# Below chunk is my data read in Info
```{r}
library(tidyr)
library(dplyr)
library(knitr)
library(rmcorr)
library(ggplot2)
library(knitr)
kb <- read.table("https://raw.githubusercontent.com/BVanSyoc/bretonData/master/data/kbAll_wAgeCats.txt", sep = "\t", header = TRUE)

rr <- read.table("https://raw.githubusercontent.com/BVanSyoc/bretonData/master/data/rrAll_wAgeCats.txt", sep = "\t", header = TRUE)


```

##Below chunk is the KB heartrate RPE Slope data and plot

```{r}
partkb <- unique(kb$Participant)
outDFkb <- data.frame()
for(i in 1:length(partkb)) {
  mod <- lm(KB_HR ~ KB_RPE, data = filter(kb, Participant == partkb[i]))
  outDFkb[i, "Participant"] <- as.numeric(partkb[i])
  outDFkb[i, "slope"] <- round(mod[1]$coefficients[2], 3)
  outDFkb[i, "intercept"] <- round(mod[1]$coefficients[1], 3)
  outDFkb[i, "rsq"] <- round(summary(mod)$r.squared, 3)
  outDFkb[i, "Sex"] <- as.character(kb$Sex[kb$Participant == partkb[i]][1])
}

summary(outDFkb)
outDFkb

mean(outDFkb$slope) #mean slope
mean(outDFkb$intercept)#mean intercept
mean(outDFkb$rsq) #mean rsquared

regkb <- lm(KB_HR ~ KB_RPE, data = kb)

perfvec <- data.frame(HR = c(60, 70, 80, 90, 100, 110, 120, 130, 140, 150, 160, 170, 180, 190, 200),
                      RPE = c(6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20))



kbreg <- ggplot() +
  geom_point(data = kb, aes(x = KB_RPE, y = KB_HR, group = Participant, color = Sex)) +
  geom_smooth(data = kb, aes(x = KB_RPE, y = KB_HR, group = Participant, color = Sex), method = "lm", se = FALSE) +
  geom_abline(intercept = mean(outDFkb$intercept), slope = mean(outDFkb$slope), color = "black", size = 1.5) +
  theme_bw()+
  geom_line(data = perfvec, aes(x = RPE, y = HR), color = "black", linetype = 5, size = 1.25) +
  labs(x ="RPE", y = "HR (bpm)") +
  scale_color_manual(values = c("blue", "red") )+
  scale_x_continuous(breaks = (6:20))+
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140, 150, 160, 170, 180, 190, 200, 210))
plot(kbreg)
```

## Chunk Below RR heartrate and RPE Slope data and plot

```{r}
partrr <- unique(rr$Participant)
outDFrr <- data.frame()
for(i in 1:length(partrr)) {
  mod <- lm(RR_HR ~ RR_RPE, data = filter(rr, Participant == partrr[i]))
  outDFrr[i, "Participant"] <- as.numeric(partrr[i])
  outDFrr[i, "slope"] <- round(mod[1]$coefficients[2], 3)
  outDFrr[i, "intercept"] <- round(mod[1]$coefficients[1], 3)
  outDFrr[i, "rsq"] <- round(summary(mod)$r.squared, 3)
  outDFrr[i, "Sex"] <- as.character(rr$Sex[rr$Participant == partrr[i]][1])
}  

summary(outDFrr)
outDFrr

mean(outDFrr$slope) #mean slope rr
mean(outDFrr$intercept) #mean intercept rr
mean(outDFrr$rsq) #mean rsquared rr

regrr <- lm(RR_HR ~ RR_RPE, data = rr)

perfvec <- data.frame(HR = c(60, 70, 80, 90, 100, 110, 120, 130, 140, 150, 160, 170, 180, 190, 200),
                      RPE = c(6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20))

rrreg <- ggplot() +
  geom_point(data = rr, aes(x = RR_RPE, y = RR_HR, group = Participant, color = Sex)) +
  geom_smooth(data = rr, aes(x = RR_RPE, y = RR_HR, group = Participant, color = Sex), method = "lm", se = FALSE) +
  geom_abline(intercept = mean(outDFrr$intercept), slope = mean(outDFrr$slope), color = "black", size = 1.5) +
  theme_bw() +
  geom_line(data = perfvec, aes(x = RPE, y = HR), color = "black", linetype = 5, size = 1.25) +
  labs(x ="RPE", y = "HR (bpm)") +
  scale_color_manual(values = c("blue", "red"))+
  scale_x_continuous(breaks = (6:20))+
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140, 150, 160, 170, 180, 190, 200, 210))
plot(rrreg)
```

# Chunk Below chankes data frame names to slope for easier understanding
```{r}
slopekb <- outDFkb
sloperr <- outDFrr
```

#Chunk below examine the sex differences for KB and RR and neither are significant
```{r}
t.test(slope ~ Sex, data = slopekb)
t.test(slope ~ Sex, data = sloperr)
```
#Chunk below examines the difference in the AFT measures between 1st and fourth quartiles to examine if VO2max may explain HR -RPE understimation - no sig difference
```{r}
#kb
t.test(slopekb$slope[slopekb$Participant %in% c(3, 6, 9, 7, 18)],
       slopekb$slope[slopekb$Participant %in% c(11, 13, 20, 21, 22)])
#rr
t.test(sloperr$slope[sloperr$Participant %in% c(2:10, 16, 18, 19, 25)],
       sloperr$slope[sloperr$Participant %in% c(1, 11:15, 17, 20:24, 26)])
```

#chunk below examines the difference in slopes in participants familiar with kb swings vs those who were not >4 rating on the 1-7 scale designated familiarity no no significant difference
```{r}
t.test(slopekb$slope[slopekb$Participant %in% c(3, 7, 9, 19, 22)],
       slopekb$slope[slopekb$Participant %in% c(1, 2, 4, 5, 6, 8, 10:18, 20, 21, 23, 24, 26)])

```

#chunk below examines the difference in slopes for KB days per week use. Those with > or equal to 1 day per week use designated from those with less often use. theres is significance.
```{r}
t.test(slopekb$slope[slopekb$Participant %in% c(3, 4, 7, 9, 10, 13, 18, 19, 22, 23, 24, 25)],
       slopekb$slope[slopekb$Participant %in% c(1, 2, 5, 6, 8, 11, 12, 14:17, 20, 21, 26)])

```

#Chunk below examines if the slopes from above are similar to the expected slope. they are significantly different. 
doesn't help explain underestimation unless we want to go with those that are familiar with kb swings reported lower rpe because they feel they are easier however they should be more intune with their body so it is not what we expect
```{r}
#test to see if slope differ from expected slpoe, top is no days per week kb use bottom is 1 or more days per week use
t.test(slopekb$slope[slopekb$Participant %in% c(3, 4, 7, 9, 10, 13, 18, 19, 22, 23, 24, 25)], mu = 10)
t.test(slopekb$slope[slopekb$Participant %in% c(1, 2, 5, 6, 8, 11, 12, 14:17, 20, 21, 26)], mu = 10)
```

#Chunk below examines the mean slope for KB and RR exercise compared to expexted slope of 10.
They are significantly different for both
```{r}
t.test(data = outDFkb, x = outDFkb$slope, mu = 10)
t.test(data = outDFrr, x = outDFrr$slope, mu = 10)
```

##AOVA Stuff

Realized I only completed a one-way ANOVA for oxygen consumption by bout for both KB an RR exercise followed by Tukey post hoc for significance. this is the reason why we discarded bout 1 in the oxygen consumption quantification because it was the only bout statistically significant between the later bouts where a plateu occured. 

Below chunk is the KB oxygen consumption by bout
```{r}
kb$Bout <- factor(kb$Bout)
kbMod <- aov(KB_VO2 ~ Bout, data = kb)

summary(kbMod)
(TukeyHSD(kbMod, conf.level = 0.95))
```

#Below is the RR oxygen consumption by bout.
bout one was sig nificantly different from all the other bouts. bout 2 was left in for the oxygen consumption mean analysis as previously discussed.
```{r}
rr$Bout <- factor(rr$Bout)
rrMod <- aov(RR_VO2 ~ Bout, data = rr)

summary(rrMod)
(TukeyHSD(rrMod, conf.level = 0.95))
```

When I realized that I only performed a one-way anova versus a repeated measured one-way anova for oxygen consumption by bout I decided to run a repeated measures ANOVA. This is where I need your input and opinion. I feel by running a repeated measures anova for oxygen consumption by bout we are essentially canceling out what we wanted to examine in the variation by participant by sepperating by bout in the normal one-way anova. I could be completly wrong and maybe we do indeed need to run a repeated measures one way anova.
I decided to run one for KB oxygen consumption by bout. doing so it said that there was no signoficant difference by bout for the kb exercise. this result differs from the one way anova by bout and what the box plot in figure one of my thesis shows visually. This is where I stopped and decided to ask for help. 

my questions are, do we need to run the repeated measures anova or are we fine with just the one way anova? if we do need to run the repeated measures anova how do we examine the statistical differences by bout like the tukey post hoc does for the normal one-way anova? when i tried to run a tukey post hoc on the repeated measures model i get the error below. I am free to chat saturday after 12:30MDT and all day sunday. If needed we can zoom meet on monday. Thanks for your help and input.

```{r}
kbvo2 <- aov(KB_VO2 ~ Bout + Error(Participant/Bout), data = kb)
summary(kbvo2)

## not running error in UseMethod("TukeyHSD") :  no applicable method for 'TukeyHSD' applied to an object of class "c('aovlist', 'listof')"
##TukeyHSD(kbvo2, conf.level = 0.95)
```

