---
title: "Summary Stats Preview"
author: "Breton Van Syoc and Emily Bean"
date: "August 6, 2019"
output:
  pdf_document: default
  html_document: default
---


```{r }
library(tidyr)
library(dplyr)
library(knitr)
library(rmcorr)
library(ggplot2)
kb <- read.table("https://raw.githubusercontent.com/BVanSyoc/bretonData/master/data/kbAll_wAgeCats.txt", sep = "\t", header = TRUE)

rr <- read.table("https://raw.githubusercontent.com/BVanSyoc/bretonData/master/data/rrAll_wAgeCats.txt", sep = "\t", header = TRUE)

```

## Energy expenditure quantification

Kettlebells

```{r }
kbsumStats <- kb %>% 
  mutate(Sex = factor(Sex),
         Participant = factor(Participant))%>% 
  group_by(Sex) %>% 
  summarize(n = length(unique(Participant)),
            avgLeanVO2 = round(mean(KB_VO2_Lean), 3),
            sdLeanVO2 = round(sd(KB_VO2_Lean), 3),
            nage18to29 = length(which(ageCats %in% "18 to 29")) / 10,
            nage30to39 = length(which(ageCats %in% "30 to 39")) / 10,
            nage40to49 = length(which(ageCats %in% "40 to 49")) / 10)
kbsumStats <- data.frame(kbsumStats)
knitr::kable(kbsumStats)
```

VO2 instead of lean VO2

```{r}
## NORMAL VO2
kbsumStats <- kb %>% 
  mutate(Sex = factor(Sex),
         Participant = factor(Participant))%>% 
  group_by(Sex) %>% 
  summarize(n = length(unique(Participant)),
            avgVO2 = round(mean(KB_VO2), 3),
            sdVO2 = round(sd(KB_VO2), 3),
            nage18to29 = length(which(ageCats %in% "18 to 29")) / 10,
            nage30to39 = length(which(ageCats %in% "30 to 39")) / 10,
            nage40to49 = length(which(ageCats %in% "40 to 49")) / 10)
kbsumStats <- data.frame(kbsumStats)
knitr::kable(kbsumStats)
```


Real runner

```{r}
rrsumStats <- rr %>% 
    mutate(Sex = factor(Sex),
         Participant = factor(Participant)) %>% 
  group_by(Sex) %>% 
  summarize(n = length(unique(Participant)),
            avgLeanVO2 = round(mean(RR_VO2_Lean), 3),
            sdLeanVO2 = round(sd(RR_VO2_Lean), 3), 
            nage18to29 = length(which(ageCats %in% "18 to 29")) / 8,
            nage30to39 = length(which(ageCats %in% "30 to 39")) / 8,
            nage40to49 = length(which(ageCats %in% "40 to 49")) / 8)
rrsumStats <- data.frame(rrsumStats)
knitr::kable(rrsumStats)
```

VO2

```{r}
rrsumStats <- rr %>% 
    mutate(Sex = factor(Sex),
         Participant = factor(Participant)) %>% 
  group_by(Sex) %>% 
  summarize(n = length(unique(Participant)),
            avgVO2 = round(mean(RR_VO2), 3),
            sdVO2 = round(sd(RR_VO2), 3), 
            nage18to29 = length(which(ageCats %in% "18 to 29")) / 8,
            nage30to39 = length(which(ageCats %in% "30 to 39")) / 8,
            nage40to49 = length(which(ageCats %in% "40 to 49")) / 8)
rrsumStats <- data.frame(rrsumStats)
knitr::kable(rrsumStats)
```
### Pearson correlation

```{r}
## KETTLEBELL
kbCor <- cor.test(kb$KB_HR, kb$KB_RPE, method = "pearson")

## RR
rrCor <- cor.test(rr$RR_HR, rr$RR_RPE, method = "pearson")
```


### ANOVA by Bout

Kettlebells

```{r}
# kettlebell ANOVA
kb$Bout <- factor(kb$Bout)
kbMod <- aov(KB_VO2 ~ Bout, data = kb)

summary(kbMod)
```

Tukey test shows significance between all bouts and bout 1
```{r}
# kettlebell Tukey test
(TukeyHSD(kbMod, conf.level = 0.95))
```

VO2 Kettle Bell

```{r}
# kettlebell ANOVA
kb$Bout <- factor(kb$Bout)
kbMod <- aov(KB_VO2 ~ Bout, data = kb)

summary(kbMod)
```

Tukey VO2 KB

```{r}
# kettlebell Tukey test
(TukeyHSD(kbMod, conf.level = 0.95))
```


Real Runner

```{r}
# real runner ANOVA
rr$Bout <- factor(rr$Bout)
rrMod <- aov(RR_VO2 ~ Bout, data = rr)

summary(rrMod)
```
Tukey test is significant between everything and bout 1, and 6, 7, 8, and bout 2
```{r}
# real runner Tukey test
(TukeyHSD(rrMod, conf.level = 0.95))
```

VO2 RR Anova

```{r}
# real runner ANOVA
rr$Bout <- factor(rr$Bout)
rrMod <- aov(RR_VO2 ~ Bout, data = rr)

summary(rrMod)
```


Tukey test V02 RR

```{r}
# real runner Tukey test
(TukeyHSD(rrMod, conf.level = 0.95))
```

##### ANOVA for HR and RPE by Bout

Kettlebell
```{r}
kbhr <- aov(KB_HR ~ Bout, data = kb)
kbrpe <- aov(KB_RPE ~ Bout, data = kb)

(TukeyHSD(kbhr))
(TukeyHSD(kbrpe))
```
```{r}
rrhr <- aov(RR_HR ~ Bout, data = rr)
rrrpe <- aov(RR_RPE ~ Bout, data = rr)

(TukeyHSD(rrhr))
(TukeyHSD(rrrpe))
```
#### above rr by bout hr and RPE

### t Test by Sex

Kettlebell

```{r}
t.test(KB_VO2_Lean ~ Sex, data = kb)
```


KBVO2 ttest

```{r}
t.test(KB_VO2 ~ Sex, data = kb)
```

Real runner
```{r}
t.test(RR_VO2_Lean ~ Sex, data = rr)
```

RR VO2 ttest
```{r}
t.test(RR_VO2 ~ Sex, data = rr)
```


### ANOVA separated by Sex

Kettlebell - males
```{r}
# ANOVA for males
kbModM <- aov(KB_VO2 ~ Bout, data = filter(kb, Sex == "M"))

summary(kbModM)
```

```{r}
# Tukey 
(TukeyHSD(kbModM, conf.level = 0.95))
```

Kettlebell - females
```{r}
# ANOVA for females
kbModF <- aov(KB_VO2 ~ Bout, data = filter(kb, Sex == "F"))

summary(kbModF)
```

```{r}
# Tukey
(TukeyHSD(kbModF, conf.level = 0.95))
```

Real runner - Males
```{r}
# ANOVA
rrModM <- aov(RR_VO2 ~ Bout, data = filter(rr, Sex == "M"))

summary(rrModM)
```

```{r}
# Tukey
(TukeyHSD(rrModM, conf.level = 0.95))
```

Real runner - Females
```{r}
# ANOVA
rrModF <- aov(RR_VO2 ~ Bout, data = filter(rr, Sex == "F"))

summary(rrModF)
```

```{r}
#Tukey
(TukeyHSD(rrModF, conf.level = 0.95))
```

### t Test by Sex, removing Bouts

Kettlebell - removing bout 1
```{r}
# filter data
kbNoBout1 <- kb %>% filter(!Bout == 1)

# t test
t.test(KB_VO2_Lean ~ Sex, data = kbNoBout1)
```

kb VO2 - removing bout one

```{r}
# filter data
kbNoBout1 <- kb %>% filter(!Bout == 1)

# t test
t.test(KB_VO2 ~ Sex, data = kbNoBout1)
```

Real runner - removing bout 1

```{r}
# filter data
rrNoBout1 <- rr %>% filter(!Bout == 1)

# t test
t.test(RR_VO2_Lean ~ Sex, data = rrNoBout1)
```

RR VO2 Removing bout 1

```{r}
# filter data
rrNoBout1 <- rr %>% filter(!Bout == 1)

# t test
t.test(RR_VO2 ~ Sex, data = rrNoBout1)
```


Kettlebells - removing bout 1 and 2

```{r}
# filter data
kbNoBout1and2 <- kb %>% filter(!Bout == c(1, 2))

t.test(KB_VO2_Lean ~ Sex, data = kbNoBout1and2)
```

Real runner - removing bout 1 and 2

```{r}
# filter data
rrNoBout1and2 <- rr %>% filter(!Bout == c(1, 2))

# t.test
t.test(RR_VO2_Lean ~ Sex, data = rrNoBout1and2)
```

## Validate HR and RPE

### repeated measures correlation

Kettlebells
```{r}
kb$Participant <- factor(kb$Participant)
kbrmc <- rmcorr(participant = Participant,
                measure1 = KB_HR,
                measure2 = KB_RPE, 
                dataset = kb,
                CIs = "analytic")
kbrmc
```

```{r}
#plot
ggplot(kb, aes(x = KB_HR, y = KB_RPE, group = Participant, color = Participant,scales = "free")) +
  geom_point() +
  geom_line(aes(y = kbrmc$model$fitted.values), linetype = 1) +
  theme_bw() +
  labs(x = "Heart Rate", y = "RPE") +
  scale_y_continuous(breaks = c(5, 7, 9, 11, 13, 15, 17, 19, 21), 
                     minor_breaks = c(6, 8, 10, 12, 14, 16, 18, 20))
```

Real runner
```{r}
rr$Participant <- factor(rr$Participant)
rrrmc <- rmcorr(participant = Participant,
                measure1 = RR_HR,
                measure2 = RR_RPE, 
                dataset = rr,
                CIs = "analytic")
rrrmc
```

```{r}
# plot
ggplot(rr, aes(x = RR_HR, y = RR_RPE, group = Participant, color = Participant,scales = "free")) +
  geom_point() +
  geom_line(aes(y = rrrmc$model$fitted.values), linetype = 1) +
  theme_bw() +
  labs(x = "Heart Rate", y = "RPE") +
  scale_y_continuous(breaks = c(5, 7, 9, 11, 13, 15, 17, 19, 21), 
                     minor_breaks = c(6, 8, 10, 12, 14, 16, 18, 20))
```

### Bland-Altman plots

Kettlebells

```{r}

# make columns for the means and differences
kbBA <- kb %>% 
  filter(!Bout == 1) %>% 
  mutate(RPE10 = KB_RPE * 10,
         avgHRRPE = (KB_HR + RPE10) / 2,
         diff = KB_HR - RPE10)

plot(x = kbBA$avgHRRPE, y = kbBA$diff, main = "Bland Altman: kettlebells")
abline(h = mean(kbBA$diff), col = "red")
abline(h = mean(kbBA$diff) + 2*sd(kbBA$diff), col = "blue")
abline(h = mean(kbBA$diff) - 2*sd(kbBA$diff), col = "blue")
text(x = 100, y = 80, label = paste0("mean: ", round(mean(kbBA$diff), 2)))
text(x = 110, y = 0, labels = paste0("CIHigh: ", round(mean(kbBA$diff) + 2*sd(kbBA$diff), 2)))
text(x = 110, y = -10, labels = paste0("CILow:", round(mean(kbBA$diff) - 2*sd(kbBA$diff), 2)))

## ggplot
ggplot(data = kbBA, aes(x = avgHRRPE, y = diff)) +
  geom_point() +
  geom_hline(yintercept = mean(kbBA$diff), col = "red") +
  geom_hline(yintercept = mean(kbBA$diff) + 2*sd(kbBA$diff), col = "blue") +
  geom_hline(yintercept = mean(kbBA$diff) - 2*sd(kbBA$diff), col = "blue") +
  theme_classic() +
  labs(x = "Mean HR (bpm)", y = "Difference HR (bpm)") +
  geom_text(aes(x = 100, y = 14 ,
                label = paste0("mean: ", round(mean(kbBA$diff), 2))), 
            size = 3.5)+
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140, 150, 160, 170, 180, 190, 200, 210))
  #geom_text(aes(x = 100, y = 0, label = paste0("SD: ", round(mean(kbBA$diff) - 2*sd(kbBA$diff), 2),
                                        #        ", ", #round(mean(kbBA$diff) + 2*sd(kbBA$diff), 2))),
            #size= 3.5)
  
```


Real runner
```{r}
rrBA <- rr %>% 
  mutate(RPE10 = RR_RPE * 10,
         avgHRRPE = (RR_HR + RPE10) /2,
         diff = RR_HR - RPE10)

plot(x = rrBA$avgHRRPE, y = rrBA$diff, main = "Bland Altman: real runner")
abline(h = mean(rrBA$diff), col = "red")
abline(h = mean(rrBA$diff) + 2*sd(rrBA$diff), col = "blue")
abline(h = mean(rrBA$diff) - 2*sd(rrBA$diff), col = "blue")
text(x = 90, y = 90, label = paste0("mean: ", round(mean(rrBA$diff), 2)))
text(x = 90, y = 0, labels = paste0("CIHigh: ", round(mean(rrBA$diff) + 2*sd(rrBA$diff), 2)))
text(x = 90, y = -10, labels = paste0("CILow:", round(mean(rrBA$diff) - 2*sd(rrBA$diff), 2)))

## ggplot
ggplot(data = rrBA, aes(x = avgHRRPE, y = diff)) +
  geom_point() +
  geom_hline(yintercept = mean(rrBA$diff), col = "red") +
  geom_hline(yintercept = mean(rrBA$diff) + 2*sd(rrBA$diff), col = "blue") +
  geom_hline(yintercept = mean(rrBA$diff) - 2*sd(rrBA$diff), col = "blue") +
  theme_classic() +
  labs(x = "Mean HR (bpm)", y = "Difference HR (bpm)") +
  geom_text(aes(x = 95, y = 14 ,
                label = paste0("mean: ", round(mean(rrBA$diff), 2))), 
            size = 3.5)+
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140, 150, 160, 170, 180, 190, 200, 210))
  #geom_text(aes(x = 95, y = 0, label = paste0("SD: ", round(mean(rrBA$diff) - 2*sd(rrBA$diff), 2),
     #                                          ", ", #round(mean(rrBA$diff) + 2*sd(rrBA$diff), 2))),
           # size= 3.5)
```

# Re-worked RMCORR figure

New plot
```{r}
## PLOT: HR on x, RPE on y, each participant with their own slope
# add rmcorr r2 = 0.7988749
# add the overall regression line

regkb <- lm(KB_RPE ~ KB_HR, data = kb)

ggplot(data = kb, aes(x = KB_HR, y = KB_RPE, group = Participant, color = Participant)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_abline(intercept = mean(outDFkb$intercept), slope = mean(outDFkb$slope), color = "black", size = 1.5) +
  theme_classic() 

```

Table showing individual r2
```{r}

# save individual slopes to a table

partkb <- unique(kb$Participant)
outDFkb <- data.frame()
for(i in 1:length(partkb)) {
  mod <- lm(KB_RPE ~ KB_HR, data = filter(kb, Participant == partkb[i]))
  outDFkb[i, "Participant"] <- as.numeric(partkb[i])
  outDFkb[i, "slope"] <- round(mod[1]$coefficients[2], 3)
  outDFkb[i, "intercept"] <- round(mod[1]$coefficients[1], 3)
  outDFkb[i, "rsq"] <- round(summary(mod)$r.squared, 3)
}

knitr::kable(outDFkb)
```

```{r}
## PLOT: HR on x, RPE on y, each participant with their own slope
# add rmcorr r = 0.7823804
# add the overall regression line

regrr <- lm(RR_RPE ~ RR_HR, data = rr)

ggplot(data = rr, aes(x = RR_HR, y = RR_RPE, group = Participant, color = Participant)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_abline(intercept = mean(outDFrr$intercept), slope = mean(outDFrr$slope), color = "black", size = 1.5) +
  theme_classic() 

```

```{r}
partrr <- unique(rr$Participant)
outDFrr <- data.frame()
for(i in 1:length(partrr)) {
  mod <- lm(RR_RPE ~ RR_HR, data = filter(rr, Participant == partrr[i]))
  outDFrr[i, "Participant"] <- as.numeric(part[i])
  outDFrr[i, "slope"] <- round(mod[1]$coefficients[2], 3)
  outDFrr[i, "intercept"] <- round(mod[1]$coefficients[1], 3)
  outDFrr[i, "rsq"] <- round(summary(mod)$r.squared, 3)
}  
knitr::kable(outDFrr)
```

