---
title: "Summary Stats Preview"
author: "Breton Van Syoc and Emily Bean"
date: "August 6, 2019"
output:
  html_document: default
---


```{r }
library(tidyr)
library(dplyr)
library(knitr)
library(rmcorr)
library(ggplot2)
kb <- read.table("https://raw.githubusercontent.com/BVanSyoc/bretonData/master/data/kbAll_wAgeCats.txt", sep = "\t", header = TRUE)

rr <- read.table("https://raw.githubusercontent.com/BVanSyoc/bretonData/master/data/rrAll_wAgeCats.txt", sep = "\t", header = TRUE)

```

## Energy expenditure quantification

Kettlebells

```{r }
kbsumStats <- kb %>% 
  mutate(Sex = factor(Sex),
         Participant = factor(Participant))%>% 
  group_by(Sex) %>% 
  summarize(n = length(unique(Participant)),
            avgLeanVO2 = round(mean(KB_VO2_Lean), 3),
            sdLeanVO2 = round(sd(KB_VO2_Lean), 3),
            nage18to29 = length(which(ageCats %in% "18 to 29")) / 10,
            nage30to39 = length(which(ageCats %in% "30 to 39")) / 10,
            nage40to49 = length(which(ageCats %in% "40 to 49")) / 10)
kbsumStats <- data.frame(kbsumStats)
knitr::kable(kbsumStats)
```

Real runner

```{r}
rrsumStats <- rr %>% 
  filter(!Bout == 1) %>% 
  mutate(Sex = factor(Sex),
         Participant = factor(Participant)) %>% 
  group_by(Sex) %>% 
  summarize(n = length(unique(Participant)),
            avgLeanVO2 = round(mean(RR_VO2_Lean), 3),
            sdLeanVO2 = round(sd(RR_VO2_Lean), 3), 
            nage18to29 = length(which(ageCats %in% "18 to 29")) / 8,
            nage30to39 = length(which(ageCats %in% "30 to 39")) / 8,
            nage40to49 = length(which(ageCats %in% "40 to 49")) / 8)
rrsumStats <- data.frame(rrsumStats)
knitr::kable(rrsumStats)
```

### ANOVA by Bout

Kettlebells

```{r}
# kettlebell ANOVA
kb$Bout <- factor(kb$Bout)
kbMod <- aov(KB_VO2_Lean ~ Bout, data = kb)

summary(kbMod)
```

Tukey test shows significance between all bouts and bout 1
```{r}
# kettlebell Tukey test
(TukeyHSD(kbMod, conf.level = 0.95))
```

Real Runner

```{r}
# real runner ANOVA
rr$Bout <- factor(rr$Bout)
rrMod <- aov(RR_VO2_Lean ~ Bout, data = rr)

summary(rrMod)
```
Tukey test is significant between everything and bout 1, and 6, 7, 8, and bout 2
```{r}
# real runner Tukey test
(TukeyHSD(rrMod, conf.level = 0.95))
```

### t Test by Sex

Kettlebell

```{r}
t.test(KB_VO2_Lean ~ Sex, data = kb)
```

Real runner
```{r}
t.test(RR_VO2_Lean ~ Sex, data = rr)
```

### ANOVA separated by Sex

Kettlebell - males
```{r}
# ANOVA for males
kbModM <- aov(KB_VO2_Lean ~ Bout, data = filter(kb, Sex == "M"))

summary(kbModM)
```

```{r}
# Tukey 
(TukeyHSD(kbModM, conf.level = 0.95))
```

Kettlebell - females
```{r}
# ANOVA for females
kbModF <- aov(KB_VO2_Lean ~ Bout, data = filter(kb, Sex == "F"))

summary(kbModF)
```

```{r}
# Tukey
(TukeyHSD(kbModF, conf.level = 0.95))
```

Real runner - Males
```{r}
# ANOVA
rrModM <- aov(RR_VO2_Lean ~ Bout, data = filter(rr, Sex == "M"))

summary(rrModM)
```

```{r}
# Tukey
(TukeyHSD(rrModM, conf.level = 0.95))
```

Real runner - Females
```{r}
# ANOVA
rrModF <- aov(RR_VO2_Lean ~ Bout, data = filter(rr, Sex == "F"))

summary(rrModF)
```

```{r}
#Tukey
(TukeyHSD(rrModF, conf.level = 0.95))
```

### t Test by Sex, removing Bouts

Kettlebell - removing bout 1
```{r}
# filter data
kbNoBout1 <- kb %>% filter(!Bout == 1)

# t test
t.test(KB_VO2_Lean ~ Sex, data = kbNoBout1)
```

Real runner - removing bout 1

```{r}
# filter data
rrNoBout1 <- rr %>% filter(!Bout == 1)

# t test
t.test(RR_VO2_Lean ~ Sex, data = rrNoBout1)
```

Kettlebells - removing bout 1 and 2

```{r}
# filter data
kbNoBout1and2 <- kb %>% filter(!Bout == c(1, 2))

t.test(KB_VO2_Lean ~ Sex, data = kbNoBout1and2)
```

Real runner - removing bout 1 and 2

```{r}
# filter data
rrNoBout1and2 <- rr %>% filter(!Bout == c(1, 2))

# t.test
t.test(RR_VO2_Lean ~ Sex, data = rrNoBout1and2)
```

## Validate HR and RPE

### repeated measures correlation

Kettlebells
```{r}
kb$Participant <- factor(kb$Participant)
kbrmc <- rmcorr(participant = Participant,
                measure1 = KB_HR,
                measure2 = KB_RPE, 
                dataset = kb,
                CIs = "analytic")
kbrmc
```

```{r}
#plot
ggplot(kb, aes(x = KB_HR, y = KB_RPE, group = Participant, color = Participant, scales = "free")) +
  geom_point() +
  geom_line(aes(y = kbrmc$model$fitted.values), linetype = 1) +
  theme_bw()
```

Real runner
```{r}
rr$Participant <- factor(rr$Participant)
rrrmc <- rmcorr(participant = Participant,
                measure1 = RR_HR,
                measure2 = RR_RPE, 
                dataset = rr,
                CIs = "analytic")
rrrmc
```

```{r}
# plot
ggplot(rr, aes(x = RR_HR, y = RR_RPE, group = Participant, color = Participant, scales = "free")) +
  geom_point() +
  geom_line(aes(y = rrrmc$model$fitted.values), linetype = 1) +
  theme_bw() 
```

### Bland-Altman plots

Kettlebells

```{r}

# make columns for the means and differences
kbBA <- kb %>% 
  mutate(RPE10 = KB_RPE * 10,
         avgHRRPE = (KB_HR + RPE10) / 2,
         diff = KB_HR - RPE10)

plot(x = kbBA$avgHRRPE, y = kbBA$diff, main = "Bland Altman: kettlebells")
abline(h = mean(kbBA$diff), col = "red")
abline(h = mean(kbBA$diff) + 2*sd(kbBA$diff), col = "blue")
abline(h = mean(kbBA$diff) - 2*sd(kbBA$diff), col = "blue")
text(x = 100, y = 80, label = paste0("mean: ", round(mean(kbBA$diff), 2)))
text(x = 110, y = 0, labels = paste0("CIHigh: ", round(mean(kbBA$diff) + 2*sd(kbBA$diff), 2)))
text(x = 110, y = -10, labels = paste0("CILow:", round(mean(kbBA$diff) - 2*sd(kbBA$diff), 2)))
```


Real runner
```{r}
rrBA <- rr %>% 
  mutate(RPE10 = RR_RPE * 10,
         avgHRRPE = (RR_HR + RPE10) /2,
         diff = RR_HR - RPE10)

plot(x = rrBA$avgHRRPE, y = rrBA$diff, main = "Bland Altman: real runner")
abline(h = mean(rrBA$diff), col = "red")
abline(h = mean(rrBA$diff) + 2*sd(rrBA$diff), col = "blue")
abline(h = mean(rrBA$diff) - 2*sd(rrBA$diff), col = "blue")
text(x = 90, y = 90, label = paste0("mean: ", round(mean(rrBA$diff), 2)))
text(x = 90, y = 0, labels = paste0("CIHigh: ", round(mean(rrBA$diff) + 2*sd(rrBA$diff), 2)))
text(x = 90, y = -10, labels = paste0("CILow:", round(mean(rrBA$diff) - 2*sd(rrBA$diff), 2)))
```

